### 2023-01-05 overview.qmd

# Overview of indicators {.unnumbered}

*Lacking content* The idea of this page is to give a quick overview of the available indicators, and to possibly present the Norwegian fact sheets for each. For now this page is just an example what it could look like.

## Areal av isbreer / Glacier area {.unnumbered}

This indicator reflects the relative loss of glacial area between tha last climatic normal persion and today.

```{r, echo=F}
source("indicators/indicatorMetaData/Areal av isbreer.R")
knitr::kable(t(out), 
            format = "html",
            caption = paste("Fakatark for tilstandindikatoren ", 
                        out$Indikator, col.names = NULL))
```

## Indicator X {.unnumbered}

And this indicator is about...